An Open-Source Implementation of the CMPS Algorithm for Assessing  Similarity of Bullet Signatures

### 2021-06-02 CMPSpaper_writing/CMPSpaper_writing.Rmd
update parameters for plotting

\wj{revised the following}
Since `npeaks.set = c(5,3,1)`, `cmps_with_multi_scale` uses the multi segment lengths strategy. On the other hand, `cmps_without_multi_scale` doesn't use the multi segment lengths strategy, and each basis segment finds 5 candidate positions due to `npeaks.set = 5`. Since `include = "full_result"` the outputs of `extract_feature_cmps(...)` contain everything listed in the previous section. Moreover, as mentioned before, multi segment lengths strategy provides a more conservative ways of finding candidate positions, so the `CMPS.score` of `cmps_with_multi_scale` is 9, while that of `cmps_without_multi_scale` is 12.
\wj{revised end}

@@

```{r hamby252-data, eval=TRUE, echo=FALSE}
# for hamby 252
N <- 3
CMPS_hamby252_results <- list()
CMPS_hamby252_results$span1 <- list(0.75, 0.25, 0.25)
CMPS_hamby252_results$signame <-
  list("sigs75", "sigs25", "sigs25_1062")
CMPS_hamby252_results$npeaks.set <-
  list(c(5, 3, 1),
       c(5, 3, 1),
       c(10, 6, 2))
CMPS_hamby252_results$seg_length <- list(50, 50, 50)
CMPS_hamby252_results$Tx <- list(25, 25, 25)
CMPS_hamby252_results$titlee <- list()
CMPS_hamby252_results$filename <- list()
for (i in 1:N) {
  CMPS_hamby252_results$titlee[[i]] <-
    paste0(
      "npeaks.set=c(",
      paste(CMPS_hamby252_results$npeaks.set[[i]], collapse = ","),
      ")",
      ", span1=",
      CMPS_hamby252_results$span1[[i]],
      ", span2=0.03, len=",
      CMPS_hamby252_results$seg_length[[i]],
      ", Tx=",
      CMPS_hamby252_results$Tx[[i]]
    )
  CMPS_hamby252_results$filename[[i]] <- 
    paste(
      "hamby252",
      CMPS_hamby252_results$span1[[i]]*100,
      paste(CMPS_hamby252_results$npeaks.set[[i]], collapse = "-"),
      CMPS_hamby252_results$seg_length[[i]],
      CMPS_hamby252_results$Tx[[i]],
      sep = "_"
    )
}
CMPS_hamby252_results$cmps.table <- list()
CMPS_hamby252_results$plot <- list()

@@

@@

  hamby252.plot.list[[1]] <- hamby252.cmps %>% ggplot() +
    geom_histogram(aes(x = cmps.max.m,
                       fill = as.factor(type_truth)), binwidth = 1) +
    labs(
      fill = "Comparison Type",
      x = expression(CMPS[max]),
      title = expression(paste("Hamby252 - ", CMPS[max], " Distribution")),
      subtitle = titlee
    ) +
    scale_x_continuous(breaks = seq(0, 27, 1)) +
    theme_bw() +
    theme(panel.grid.minor = element_blank())

  hamby252.plot.list[[2]] <- hamby252.cmps %>% ggplot() +
    geom_histogram(aes(x = cmps.maxbar.m,
                       fill = as.factor(type_truth)), binwidth = 1) +
    labs(
      x = expression(bar(CMPS)[max]),
      fill = "Comparison Type",
      title = expression(paste(
        "Hamby252 - ", bar(CMPS)[max], " Distribution"
      )),
      subtitle = titlee
    ) +
    scale_x_continuous(breaks = seq(0, 24, 1)) +
    theme_bw() +
    theme(panel.grid.minor = element_blank())

  plot <- ggarrange(plotlist = hamby252.plot.list,
                    nrow = 1,
                    ncol = 2)
  CMPS_hamby252_results$plot[[i]] <- plot
}

@@

```{r, eval=TRUE, echo=TRUE, fig.height=3}
CMPS_hamby252_results$plot[[1]]
```

```{r hamby44-data, eval=TRUE, echo=FALSE}
N <- 3
CMPS_hamby44_results <- list()
CMPS_hamby44_results$span1 <- list(0.25, 0.25, 0.75)
CMPS_hamby44_results$signame <-
  list("sigs25_531", "sigs25_1061", "sigs75_531")
CMPS_hamby44_results$npeaks.set <-
  list(c(5, 3, 1),
       c(10, 6, 1),
       c(5, 3, 1))
CMPS_hamby44_results$seg_length <- as.list(rep(61, 3))
CMPS_hamby44_results$Tx <- as.list(rep(30, 3))
CMPS_hamby44_results$titlee <- list()
CMPS_hamby44_results$filename <- list()
for (i in 1:N) {
  CMPS_hamby44_results$titlee[[i]] <-
    paste0(
      "npeaks.set=c(",
      paste(CMPS_hamby44_results$npeaks.set[[i]], collapse = ","),
      ")",
      ", span1=",
      CMPS_hamby44_results$span1[[i]],
      ", span2=0.03, len=",
      CMPS_hamby44_results$seg_length[[i]],
      ", Tx=",
      CMPS_hamby44_results$Tx[[i]]
    )
  CMPS_hamby44_results$filename[[i]] <- 
    paste(
      "hamby44",
      CMPS_hamby44_results$span1[[i]]*100,
      paste(CMPS_hamby44_results$npeaks.set[[i]], collapse = "-"),
      CMPS_hamby44_results$seg_length[[i]],
      CMPS_hamby44_results$Tx[[i]],
      sep = "_"
    )
}
CMPS_hamby44_results$cmps.table <- list()
CMPS_hamby44_results$plot <- list()
```

@@

Figure 2 shows the distribution of $\mathrm{CMPS_{max}}$ and $\mathrm{\overline{CMPS}_{max}}$ after we applied the CMPS algorithm to Hamby set 252. Although the CMPS scores we found here are not exactly the same as those presented in @cmps since we don't have the exact parameters and functions used by the researchers, the results we obtained are promising. For both $\mathrm{CMPS_{max}}$ and $\mathrm{\overline{CMPS}_{max}}$ we can observe a clear separation between the known matching comparisons (KM) and known non-matching comparisons (KNM). Table x shows some summary statistics.

@@

Note that for Hamby set 44, one unit represents 1.5625 \textmu m so setting `seg_length = 50` makes each basis segment have length 78.125 \textmu m. And `Tx = 25` makes the tolerance zone be $\pm 39.0625$ \textmu m. This is to mimic the parameters used in the [original paper].

<!-- this is figure 2 -->
```{r, echo=FALSE, out.width="400px", fig.cap="Distribution of $\\mathrm{CMPS_{max}}$ and $\\mathrm{\\overline{CMPS}_{max}}$ for Hamby 252; removed outliers; seg\\_length = 50, Tx = 25, npeaks.set = c(5,3,1) "}
# [4] "span1=0.25, span2=0.03, rm_outliers = TRUE, 5025, 531"  
knitr::include_graphics("img/hamby252_v1.png", dpi = 100)
```

table 1:

```{r, echo=FALSE, warning=FALSE}
# rtt <- readRDS("summary_table.rds")
# colnames(rtt) <- 
rtt <- data.frame(parameters = c("span1=0.25, (5, 3, 1)", "span1=0.25, (10, 6, 1)", "span1=0.75, (5, 3, 1)"),
                  p1 = c(7,9,6),
                  p2 = c(10, 13, 11),
                  p3 = c(3.6, 4.833, 3.167),
                  p4 = c(6.667, 9.167, 6.6))
colnames(rtt) <- c("Parameters",
                   "$\\mathrm{CMPS_{max}}$ KNM's max",
                   "$\\mathrm{CMPS_{max}}$ KM's min",
                   "$\\mathrm{\\overline{CMPS}_{max}}$ KNM's max",
                   "$\\mathrm{\\overline{CMPS}_{max}}$ KM's min")

@@

Figure 3 shows the results for Hamby set 44. Again, we are able to see a clear separation between the KM comparisons and the KNM comparisons, but it would be better if the separation can be further enlarged. Table 1 shows some summary statistics for Hamby set 44.

@@

```{r, echo=FALSE, out.width="400px", fig.cap="Distribution of $\\mathrm{CMPS_{max}}$ and $\\mathrm{\\overline{CMPS}_{max}}$ for Hamby 44; removed outliers; seg\\_length = 61, Tx = 30, npeaks.set = c(5,3,1), span1=0.25"}
# [1] "span1=0.25, span2=0.03, rm_outliers"
knitr::include_graphics("img/hamby44_v1.png", dpi = 100)
```

Figure 4 shows another Hamby set 44 results, but the parameter `npeaks.set` is now `c(10, 6, 1)`. We can see that the separation between KM comparisons and KNM comparisons is increased. This shows that the parameters used in the CMPS algorithm can affect the CMPS results, and an open-source implementation of the algorithm will facilitate efforts of cross-validating these parameters.

@@

```{r, echo=FALSE, out.width="400px", fig.cap="Distribution of $\\mathrm{CMPS_{max}}$ and $\\mathrm{\\overline{CMPS}_{max}}$ for Hamby 44; removed outliers; seg\\_length = 61, Tx = 30, npeaks.set = c(10,6,1), span1=0.25 "}
# [5] "span1=0.25, span2=0.03, rm_outliers, c(10,6,1)"
knitr::include_graphics("img/hamby44_c1_v1.png", dpi = 100)
```

For the results above we used `span1=0.25` in the function `cc_get_signature` when computing the signatures; however, in this example we used `span1=0.75` and changed the characteristic of those signatures. As we can see in figure 5, this change also affects the CMPS results.

@@

```{r, echo=FALSE, out.width="400px", fig.cap="Distribution of $\\mathrm{CMPS_{max}}$ and $\\mathrm{\\overline{CMPS}_{max}}$ for Hamby 44; removed outliers; seg\\_length = 61, Tx = 30, npeaks.set = c(5,3,1), span1=0.75"}
# [4] "span1=0.75, span2=0.03, rm_outliers"  
knitr::include_graphics("img/hamby44_c2_v1.png", dpi = 100)
```

>>>>

\wj{revised the following}
Since `npeaks.set = c(5,3,1)`, `cmps_with_multi_scale` uses the multi segment lengths strategy. On the other hand, `cmps_without_multi_scale` doesn't use the multi segment lengths strategy, and each basis segment finds 5 candidate positions due to `npeaks.set = 5`. Since `include = "full_result"` the outputs of `extract_feature_cmps(...)` contain everything listed in the previous section. Moreover, \wj{XXX as mentioned before, }multi segment lengths strategy provides a more conservative ways of finding candidate positions, so the `CMPS.score` of `cmps_with_multi_scale` is 9, while that of `cmps_without_multi_scale` is 12. \wj{XXX how should we address the advantage of multi segment lengths strategy}
\wj{revised end}

@@

```{r hamby252-data, eval=TRUE, echo=FALSE}
# for hamby 252
N <- 3
CMPS_hamby252_results <- list()
CMPS_hamby252_results$span1 <- list(0.75, 0.25, 0.25)
CMPS_hamby252_results$signame <-
  list("sigs75", "sigs25", "sigs25_1062")
CMPS_hamby252_results$npeaks.set <-
  list(c(5, 3, 1),
       c(5, 3, 1),
       c(10, 6, 2))
CMPS_hamby252_results$seg_length <- list(50, 50, 50)
CMPS_hamby252_results$Tx <- list(25, 25, 25)
CMPS_hamby252_results$titlee <- list()
CMPS_hamby252_results$filename <- list()
for (i in 1:N) {
  CMPS_hamby252_results$titlee[[i]] <-
    paste0(
      "npeaks.set=c(",
      paste(CMPS_hamby252_results$npeaks.set[[i]], collapse = ","),
      ")",
      ", len=",
      CMPS_hamby252_results$seg_length[[i]],
       ", Tx=",
      CMPS_hamby252_results$Tx[[i]],
      ", \nspan1=",
      CMPS_hamby252_results$span1[[i]],
      ", span2=0.03"
    )
  CMPS_hamby252_results$filename[[i]] <- 
    paste(
      "hamby252",
      CMPS_hamby252_results$span1[[i]]*100,
      paste(CMPS_hamby252_results$npeaks.set[[i]], collapse = "-"),
      CMPS_hamby252_results$seg_length[[i]],
      CMPS_hamby252_results$Tx[[i]],
      sep = "_"
    )
}
CMPS_hamby252_results$cmps.table <- list()
CMPS_hamby252_results$plot <- list()

@@

com.title252 <- expression(paste(
  "Hamby 252 - ", CMPS[max], " and ", bar(CMPS)[max], " Distribution"
))

@@

  hamby252.plot.list[[1]] <- hamby252.cmps %>% ggplot() +
    geom_histogram(aes(x = cmps.max.m,
                       fill = as.factor(type_truth)), binwidth = 1) +
    labs(
      fill = "Comparison Type",
      x = expression(CMPS[max]),
      # title = expression(paste("Hamby252 - ", CMPS[max], " Distribution")),
      subtitle = titlee
    ) +
    scale_x_continuous(breaks = seq(0, 27, 1)) +
    theme_bw() +
    theme(panel.grid.minor = element_blank()) +
    font("x.text", size = 6)

  hamby252.plot.list[[2]] <- hamby252.cmps %>% ggplot() +
    geom_histogram(aes(x = cmps.maxbar.m,
                       fill = as.factor(type_truth)), binwidth = 1) +
    labs(
      x = expression(bar(CMPS)[max]),
      fill = "Comparison Type",
      # title = expression(paste(
      #   "Hamby252 - ", bar(CMPS)[max], " Distribution"
      # )),
      subtitle = titlee
    ) +
    scale_x_continuous(breaks = seq(0, 24, 1)) +
    theme_bw() +
    theme(panel.grid.minor = element_blank())

  plot <- ggarrange(plotlist = hamby252.plot.list,
                    nrow = 1,
                    ncol = 2,
                    common.legend = TRUE, legend = "bottom")
  plot <- annotate_figure(plot, 
                          top = text_grob(com.title252))
  CMPS_hamby252_results$plot[[i]] <- plot
}

@@

```{r hamby44-data, eval=TRUE, echo=FALSE}
N <- 3
CMPS_hamby44_results <- list()
CMPS_hamby44_results$span1 <- list(0.25, 0.25, 0.75)
CMPS_hamby44_results$signame <-
  list("sigs25_531", "sigs25_1061", "sigs75_531")
CMPS_hamby44_results$npeaks.set <-
  list(c(5, 3, 1),
       c(10, 6, 1),
       c(5, 3, 1))
CMPS_hamby44_results$seg_length <- as.list(rep(61, 3))
CMPS_hamby44_results$Tx <- as.list(rep(30, 3))
CMPS_hamby44_results$titlee <- list()
CMPS_hamby44_results$filename <- list()
for (i in 1:N) {
  CMPS_hamby44_results$titlee[[i]] <-
    paste0(
      "npeaks.set=c(",
      paste(CMPS_hamby44_results$npeaks.set[[i]], collapse = ","),
      ")",
      ", len=",
      CMPS_hamby44_results$seg_length[[i]],
       ", Tx=",
      CMPS_hamby44_results$Tx[[i]],
      ", \nspan1=",
      CMPS_hamby44_results$span1[[i]],
      ", span2=0.03"
    )
  CMPS_hamby44_results$filename[[i]] <- 
    paste(
      "hamby44",
      CMPS_hamby44_results$span1[[i]]*100,
      paste(CMPS_hamby44_results$npeaks.set[[i]], collapse = "-"),
      CMPS_hamby44_results$seg_length[[i]],
      CMPS_hamby44_results$Tx[[i]],
      sep = "_"
    )
}
CMPS_hamby44_results$cmps.table <- list()
CMPS_hamby44_results$plot <- list()

for(i in 1:N){
  CMPS_hamby44_results$cmps.table[[i]] <- 
  read.csv(file = paste("./data/", 
                        CMPS_hamby44_results$filename[[i]], 
                        ".csv",
                        sep = ""))
}

com.title44 <- expression(paste(
  "Hamby 44 - ", CMPS[max], " and ", bar(CMPS)[max], " Distribution"
))

# generate plots
for (i in 1:3) {
  hamby44.cmps <- CMPS_hamby44_results$cmps.table[[i]]

  hamby44.plot.list <- list()
  titlee <- CMPS_hamby44_results$titlee[[i]]

  hamby44.plot.list[[1]] <- hamby44.cmps %>% ggplot() +
    geom_histogram(aes(x = cmps.max.m,
                       fill = as.factor(type_truth)), binwidth = 1) +
    labs(
      fill = "Comparison Type",
      x = expression(CMPS[max]),
      # title = expression(paste("Hamby252 - ", CMPS[max], " Distribution")),
      subtitle = titlee
    ) +
    scale_x_continuous(breaks = seq(0, 27, 1)) +
    theme_bw() +
    theme(panel.grid.minor = element_blank()) +
    font("x.text", size = 6)

  hamby44.plot.list[[2]] <- hamby44.cmps %>% ggplot() +
    geom_histogram(aes(x = cmps.maxbar.m,
                       fill = as.factor(type_truth)), binwidth = 1) +
    labs(
      x = expression(bar(CMPS)[max]),
      fill = "Comparison Type",
      # title = expression(paste(
      #   "Hamby252 - ", bar(CMPS)[max], " Distribution"
      # )),
      subtitle = titlee
    ) +
    scale_x_continuous(breaks = seq(0, 24, 1)) +
    theme_bw() +
    theme(panel.grid.minor = element_blank())

  plot <- ggarrange(plotlist = hamby44.plot.list,
                    nrow = 1,
                    ncol = 2,
                    common.legend = TRUE, legend = "bottom")
  plot <- annotate_figure(plot, 
                          top = text_grob(com.title44))
  CMPS_hamby44_results$plot[[i]] <- plot
}
```

@@

\autoref{fig:result1_252} shows the distribution of $\mathrm{CMPS_{max}}$ and $\mathrm{\overline{CMPS}_{max}}$ after we applied the CMPS algorithm to Hamby set 252. Although the CMPS scores we found here are not exactly the same as those presented in @cmps since we don't have the exact parameters and functions used by the researchers, the results we obtained are promising. For both $\mathrm{CMPS_{max}}$ and $\mathrm{\overline{CMPS}_{max}}$ we can observe a clear separation between the known matching comparisons (KM) and known non-matching comparisons (KNM). \autoref{tab:table1} shows some summary statistics.

@@

Note that for Hamby set 44, one unit represents 1.5625 \textmu m so setting `seg_length = 50` makes each basis segment have length 78.125 \textmu m. And `Tx = 25` makes the tolerance zone be $\pm 39.0625$ \textmu m. This is to mimic the parameters used in @cmps.

<!-- this is figure 2 -->
```{r result1_252, echo=FALSE, out.width="400px", fig.cap="Distribution of $\\mathrm{CMPS_{max}}$ and $\\mathrm{\\overline{CMPS}_{max}}$ for Hamby 252; removed outliers; seg\\_length = 50, Tx = 25, npeaks.set = c(5,3,1) "}
# [4] "span1=0.25, span2=0.03, rm_outliers = TRUE, 5025, 531"  
# knitr::include_graphics("img/hamby252_v1.png", dpi = 100)
CMPS_hamby252_results$plot[[1]]
```

Table 1:

```{r table1, echo=FALSE, warning=FALSE}
# rtt <- readRDS("summary_table.rds")
# colnames(rtt) <- 
rtt <- data.frame(parameters = c("span1=0.25, (5, 3, 1)", "span1=0.25, (10, 6, 1)", "span1=0.75, (5, 3, 1)"),
                  p1 = c(7,9,6),
                  p2 = c(10, 13, 11),
                  p3 = c(3.6, 4.833, 3.167),
                  p4 = c(6.667, 9.167, 6.6))
colnames(rtt) <- c("Parameters",
                   "$\\mathrm{CMPS_{max}}$ KNM's max",
                   "$\\mathrm{CMPS_{max}}$ KM's min",
                   "$\\mathrm{\\overline{CMPS}_{max}}$ KNM's max",
                   "$\\mathrm{\\overline{CMPS}_{max}}$ KM's min")

@@

\autoref{fig:result1_44} shows the results for Hamby set 44. Again, we are able to see a clear separation between the KM comparisons and the KNM comparisons, but it would be better if the separation can be further enlarged. \wj{Table x} shows some summary statistics for Hamby set 44.

@@

```{r result1_44, echo=FALSE, out.width="400px", fig.cap="Distribution of $\\mathrm{CMPS_{max}}$ and $\\mathrm{\\overline{CMPS}_{max}}$ for Hamby 44; removed outliers; seg\\_length = 61, Tx = 30, npeaks.set = c(5,3,1), span1=0.25"}
# [1] "span1=0.25, span2=0.03, rm_outliers"
# knitr::include_graphics("img/hamby44_v1.png", dpi = 100)
CMPS_hamby44_results$plot[[1]]
```

\autoref{fig:result2_44} shows another Hamby set 44 results, but the parameter `npeaks.set` is now `c(10, 6, 1)`. We can see that the separation between KM comparisons and KNM comparisons is increased. This shows that the parameters used in the CMPS algorithm can affect the CMPS results, and an open-source implementation of the algorithm will facilitate efforts of cross-validating these parameters.

@@

```{r result2_44, echo=FALSE, out.width="400px", fig.cap="Distribution of $\\mathrm{CMPS_{max}}$ and $\\mathrm{\\overline{CMPS}_{max}}$ for Hamby 44; removed outliers; seg\\_length = 61, Tx = 30, npeaks.set = c(10,6,1), span1=0.25 "}
# [5] "span1=0.25, span2=0.03, rm_outliers, c(10,6,1)"
# knitr::include_graphics("img/hamby44_c1_v1.png", dpi = 100)
CMPS_hamby44_results$plot[[2]]
```

For the results above we used `span1=0.25` in the function `cc_get_signature` when computing the signatures; however, in this example we used `span1=0.75` and changed the characteristic of those signatures. As we can see in \autoref{fig:result3_44}, this change also affects the CMPS results.

@@

```{r result3_44, echo=FALSE, out.width="400px", fig.cap="Distribution of $\\mathrm{CMPS_{max}}$ and $\\mathrm{\\overline{CMPS}_{max}}$ for Hamby 44; removed outliers; seg\\_length = 61, Tx = 30, npeaks.set = c(5,3,1), span1=0.75"}
# [4] "span1=0.75, span2=0.03, rm_outliers"  
# knitr::include_graphics("img/hamby44_c2_v1.png", dpi = 100)
CMPS_hamby44_results$plot[[3]]
```
